\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\newcommand*{\Perm}[2]{{}^{#1}\!P_{#2}}%
\title{Pre-test practice recitation}
\date{\today}
\author{CS161}
\begin{document}
\maketitle


\section{General Instructions}
Refer to the PDF in the class code (july-11/resources): \\
\href{https://github.com/csu161/Class-code/blob/master/july-11/resources/inductive-proof-examples.pdf}{Inductive Proof Examples}


\section{Counting}
\begin{enumerate}
\item How many positive integers less than 1000
  \begin{enumerate}
  \item have exactly three digits?
  \item have an odd number of digits?
  \item have at least one digit equal to 9?
  \item have no odd digits?
  \item have two consecutive digits equal to 5?
  \item are palindromes?
  \end{enumerate}
\item How many ways are there to choose a dozen donuts from 20 varieties
  \begin{enumerate}
  \item if there are no two donuts of the same variety?
  \item if all donuts are of the same variety?
  \item if there are no restrictions?
  \item if there are at least two varieties?
  \item if there must be at least six blueberry donuts?
  \end{enumerate}
\item In how many different ways can five elements be selected in order from a set with three elements when repetition is allowed?
\item How many strings of six letters are there?
  \begin{enumerate}
  \item if only lowercase is allowed?
  \item if any case is allowed?
  \end{enumerate}
\item In how many ways can you stack 7 different books, so that a specific book B is on the third place?
\item In how many ways can you take 3 marbles out of a box with 15 different marbles?
\item In how many ways can you take 5 cards, with at least 2 aces, out of a deck of 52 cards?
\item Find $n$ if
  \begin{enumerate}
  \item $\Perm{n}{2} = 110$
  \item $\Perm{n}{n} = 5040$
  \item $\Perm{n}{4} = 12 * \Perm{n}{2}$
  \end{enumerate}
\item Find $n$ if
  \begin{enumerate}
  \item ${n \choose 2} = 45$
  \item ${n \choose 3} = \Perm{n}{2}$
  \item ${n \choose 5} = {n \choose 2}$
  \end{enumerate}
\item If the numbers from 1 to 1000 are written out on a piece of paper, how many 9's are on that paper?
\end{enumerate}



\section{Proofs}
\begin{enumerate}
\item Prove that for all positive integers, $n$:
\[ \frac{2}{3} + \frac{2}{9} + \frac{2}{27} + ... + \frac{2}{3^n} = 1 - \frac{1}{3^n} \]
\item Which amounts of postage can you make using 5 and 9 cent stamps?
\item Find $f(1)$, $f(2)$, $f(3)$, and $f(4)$ if $f(n)$ is defined recursively by $f(0) = 1$ and for $n = 0, 1, 2, ...$
  \begin{enumerate}
  \item $f(n + 1) = f(n) + 2$
  \item $f(n + 1) = 3f(n)$
  \item $f(n + 1) = 2^{f(n)}$
  \item $f(n + 1) = f(n)^2 + f(n) + 1$
  \end{enumerate}
\item Let $P(n)$ be the statement that
\[ 1 + \frac{1}{4} + \frac{1}{9} + ... + \frac{1}{n^2} < 2 - \frac{1}{n} \]
where $n$ is an integer greater than 1.
  \begin{enumerate}
  \item What is the statement $P(2)$?
  \item Show that $P(2)$ is true, completing the basis step of the proof.
  \item What is the \textit{inductive hypothesis}?
  \item What do you need to prove in the inductive step?
  \item Complete the inductive step.
  \item Explain why these steps show that this inequality is true for all $n$ where $n$ is an integer greater than 1.
  \end{enumerate}
\item Prove by induction that:
\[ 1 + 3 + 5 + ... + (2n - 1) = n^2 \]
\end{enumerate}

\end{document}
